Paper Analysis

### 2025-04-18 Paper_WriteUp/Paper_Plots.qmd
consistent model names and fixing xlim on fig4 right

#read in models
model_17_19 = readRDS(here("results_revision", "lichess1700-1900", "all_rated_bullet_model_prev_n1.RDS"))
model_20_22 = readRDS(here("results_revision", "lichess2000-2200", "all_rated_bullet_model_prev_n1.RDS"))
model_23_25 = readRDS(here("results_revision", "lichess2300-2500", "all_rated_bullet_model_prev_n1.RDS"))
model_gm = readRDS(here("results_revision", "lichessGrandmasters", "all_rated_bullet_model_prev_n1.RDS"))

#get global effects of each
pars_17_19 <- model_17_19$draws(c("mu_beta", "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "1700-1900")

pars_20_22 = model_20_22$draws(c("mu_beta", "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "2000-2200")

pars_23_25 = model_23_25$draws(c("mu_beta", "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "2300-2500")

pars_gm <- model_gm$draws(c("mu_beta", "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "GM")

#for plotting
facet_labels <- as_labeller(c(
  mu_beta = "mu[beta]",
  gamma1 = "gamma[1]",
  gamma2 = "gamma[2]",
  sigma_1 = "sigma[1]",
  `tau[1]` = "tau[1]",
  `tau[2]` = "tau[2]",
  sigma_g1 = "sigma[g[1]]",
  sigma_g2 = "sigma[g[2]]"
), label_parsed)

comp_global <- bind_rows(pars_17_19, pars_20_22, pars_23_25, pars_gm) |> 
  ggplot(aes(x = value, y = name, color = factor(model, levels = c("1700-1900", "2000-2200", "2300-2500", "GM")))) +
  stat_pointinterval(position = position_dodge(width = 0.2)) +
  facet_wrap(~name,
             scales = "free",
             labeller = labeller(name = facet_labels)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_3col() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

@@

```{r player level parameters for 17-19}
#| eval: false

player_id <- as_tibble(bullet_17_19) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

@@

```{r figure s4 left - plot player effects}
#| eval: false
indiv_plot_17_19 <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
  stat_histinterval() +
  labs(x = "Estimated Player Effect", y = element_blank(),
       title = "1700-1900 Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  xlim(c(-0.1, 0.1)) +
  theme_single_y()

@@

## then the winner effects plot
winner_plot_gm <- player_pars |> 
  filter(param == 2) |> 
  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |>
  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
  stat_histinterval() +
  labs(x = "Estimated W/L-Effect", y = element_blank(), 
       title = "GM Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

winner_plot_gm

@@

@@

### mu_beta simulation comparison

>>>>

#read in models
fit_17_19_bullet_prev  = readRDS(here("results_revision", "lichess1700-1900", "all_rated_bullet_model_prev_n1.RDS"))
fit_20_22_bullet_prev  = readRDS(here("results_revision", "lichess2000-2200", "all_rated_bullet_model_prev_n1.RDS"))
fit_23_25_bullet_prev = readRDS(here("results_revision", "lichess2300-2500", "all_rated_bullet_model_prev_n1.RDS"))
fit_gm_bullet_prev  = readRDS(here("results_revision", "lichessGrandmasters", "all_rated_bullet_model_prev_n1.RDS"))

#read in users
bullet_17_19 = readRDS(here("results_revision", "lichess1700-1900", "users_bullet.RDS"))
bullet_20_22 = readRDS(here("results_revision", "lichess2000-2200", "users_bullet.RDS"))
bullet_23_25 = readRDS(here("results_revision", "lichess2300-2500", "users_bullet.RDS"))
bullet_gm = readRDS(here("results_revision", "lichessGrandmasters", "users_bullet.RDS"))

#get global effects of each
pars_17_19 = fit_17_19_bullet_prev $draws(c("mu_beta", "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "1700-1900")

pars_20_22 = fit_20_22_bullet_prev$draws(c("mu_beta", "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "2000-2200")

pars_23_25 = fit_23_25_bullet_prev$draws(c("mu_beta", "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "2300-2500")

pars_gm = fit_gm_bullet_prev$draws(c("mu_beta", "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "GM")

#for plotting
facet_labels = as_labeller(c(
  mu_beta = "mu[beta]",
  gamma1 = "gamma[1]",
  gamma2 = "gamma[2]",
  sigma_1 = "sigma[1]",
  `tau[1]` = "tau[1]",
  `tau[2]` = "tau[2]",
  sigma_g1 = "sigma[g[1]]",
  sigma_g2 = "sigma[g[2]]"
), label_parsed)

comp_global = bind_rows(pars_17_19, pars_20_22, pars_23_25, pars_gm) |> 
  ggplot(aes(x = value, y = name, color = factor(model, levels = c("1700-1900", "2000-2200", "2300-2500", "GM")))) +
  stat_pointinterval(position = position_dodge(width = 0.2)) +
  facet_wrap(~name,
             scales = "free",
             labeller = labeller(name = facet_labels)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_3col() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

@@

```{r player level parameters for 17-19}
#| eval: false
player_id <- as_tibble(bullet_17_19) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

@@

```{r figure s4 left - plot player effects}
#| eval: false
indiv_plot_17_19 <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
  stat_histinterval() +
  labs(x = "Estimated Player Effect", y = element_blank(),
       title = "1700-1900 Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

@@

## then the winner effects plot
winner_plot_gm <- player_pars |> 
  filter(param == 2) |> 
  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |>
  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
  stat_histinterval() +
  labs(x = "Estimated W/L-Effect", y = element_blank(), 
       title = "GM Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  xlim(c(-0.1, 0.1)) +
  theme_single_y()

@@

# Beta simulation comparison

@@

# mu_beta simulation comparison
